RTS AI: Problems and Techniques

### 2015-08-20 ECGG_RTSProblems/ECGG_RTSProblems.tex
Reshape for submission (like get rid of the figures folder)

\subsubsection{Planning}
As mentioned above, the  size of the state space in  RTS games is much
larger  than  that  of  traditional  board  games  such  as  Chess  or
Go. Additionally,  the number  of actions  that can  be executed  at a
given instant of time is  also much larger. Thus, standard adversarial
planning  approaches,  such  as  game tree  search  are  not  directly
applicable. As we  elaborate later, planning in RTS games  can be seen
as having multiple  levels of abstraction: at a  higher level, players
need long-term  planning capabilities,  in order  to develop  a strong
economy in the game; at a low level, individual units need to be moved
in coordination to  fight battles taking into account  the terrain and
the  opponent.  Techniques  that  can  address  these  large  planning
problems by either sampling, or  hierarchical decomposition do not yet
exist.

\subsubsection{Learning}
Given  the  difficulties  in  playing  by  directly  using
adversarial  planning techniques,  many  research  groups have  turned
attention to  learning techniques. We  can distinguish three  types of
learning problems in RTS games:
\begin{itemize}
\item {\em Prior learning}: How can we exploit available data, such as
  existing replays,  or information  about specific maps  for learning
  appropriate strategies before hand? A significant amount of work has
  gone in this
  direction.%, such as \cite{WeberCIG09,SynnaeveCIG11,SynnaeveAIIDE11,OntanonMSR10}.
\item  {\em In-game  learning}: How  can bots  deploy online  learning
  techniques that allow them to  improve their game play while playing
  a  game?  These  techniques  might  include  reinforcement  learning
  techniques, but  also opponent modeling.  The main problem  again is
  the fact  that the state  space is too large  and the fact  that RTS
  games are partially observable.
\item {\em  Inter-game learning}:  What can be  learned from  one game
  that can  be used  to increase  the chances of  victory in  the next
  game? Some work has used simple game-theoretical solutions to select
  amongst a  pool of  predefined strategies,  but the  general problem
  remains unsolved.
\end{itemize}

@@

\subsubsection{Uncertainty}
Adversarial planning under  uncertainty in domains of the  size of RTS
games is  still an unsolved  challenge.  In  RTS games, there  are two
main kinds  of uncertainty. First,  the game is  partially observable,
and players  cannot observe the  whole game  map (like in  Chess), but
need to scout in order to see what the opponent is doing. This type of
uncertainty   can  be   lowered  by   good  scouting,   and  knowledge
representation  (to  infer  what  is  possible  given  what  has  been
seen). However, scouting also means deliberately reducing economic progress in order to obtain information.
Second,  there is also  uncertainty arising from the  fact that
the games  are adversarial,  and a player  cannot predict  the actions
that the opponent(s)  will execute. For this type  of uncertainty, the
AI, as the human  player, can only build a sensible  model of what the
opponent is likely to do.
% [Santi:  I  commented  out  this  part,  since  I'm  not  sure  it's
% correct.  The second  type  of  uncertainty can  also  be linked  to
% data.  For example,  by analyzing  player replays,  one could  learn
% "plan recognition"  probability tables, which  can be used  to infer
% the strategy  of the opponent,  and thus, reduce the  uncertainty of
% the opponent's future  actions] Both these kinds  of uncertainty can
% be studied through probabilities, but only the first type, $P(Hidden
% | Seen)$, can be {\em easily} linked to
% data. %now for the clumsy part:
% One could  argue that the  extensional data  is the only  thing that
% matters because  it is the only  thing that has an  influence on the
% game, but in  reality, for very highly  skilled players, discovering
% the intentions of the  enemy is the key to winning.  It allows for a
% compact understanding of the game  state and development, while also
% giving additional information about the  future. Whereas a player is
% doing action  A to  follow it  up by action  B or  C depends  on its
% intention,  while the  information  about action  A  is visible  and
% information about  B or  C is  hidden, only  intention allows  for a
% choice.

\subsubsection{Spatial and Temporal Reasoning}
% TODO just a paragraph on exposing why there are temporal and spatial
% reasoning in RTS AI.
Spatial reasoning is  related to each aspect  of terrain exploitation.
It  is  involved   in  tasks  such  as  building   placement  or  base
expansion.  In the  former,  the player  needs  to carefully  consider
building  positioning into  its  own  bases to  both  protect them  by
creating  defenses or walls  against invasions  and to  avoid bad  configurations
where large units could be stuck. In base expansion, the player has to
choose good available locations to build a new base, regarding its own
position and  opponent's bases. Finally,  spatial reasoning is  key to
tactical reasoning:  players need to  decide where to place  units for
battle, favoring, for instance,  engagements when the opponent's units
are lead into a bottleneck.

@@

\subsubsection{Domain Knowledge Exploitation}
In traditional board  games such as Chess,  researchers have exploited
the  large  amounts  of  existing  domain  knowledge  to  create  good
evaluation  functions  to be  used  by  alpha-beta search  algorithms,
extensive opening books, or end-game tables. In the case of RTS games,
it  is still  unclear how  the  significantly large  amount of  domain
knowledge  (in the  forms or  strategy guides,  replays, etc.)  can be
exploited. Most  work in  this area has  focused on  two main
directions: (1) hard-coding existing strategies into bots (so that  bots only  need to
decide  which strategies  to deploy,  instead of  having to  solve the
complete  planning problem), and (2) mining large datasets
of  replays \cite{WeberCig09,synnaeve2012dataset} to automatically extract
 strategies,  trends   or  plans. % However,  StarCraft  games  are  quite  complex,  and  how  to
%automatically learn from such datasets is still an open problem.

\subsubsection{Task Decomposition}
Most existing approaches  to play RTS games work by decomposing the  problem into   a    collection   of    smaller   problems to    be   solved
independently. Specifically, a common subdivision is:
\begin{itemize}
\item {\em  Strategy}: corresponds  to the high-level  decision making
  process.  This is  the highest  level  of abstraction  for the  game
  comprehension.  Finding an  efficient  strategy or  counter-strategy
  against a given opponent is key in RTS games, and concerns the whole
  set of units a player
  owns. %If we may allow ourselves a military analogy, strategy is orders given by generals at the HQ.
\item   {\em  Tactics}:   are  the   implementation  of   the  current
  strategy.  It  implies  army and  building  positioning,  movements,
  timing, and so on. Tactics concerns a group of
  units. %They are orders given by officers on the battlefield.
\item {\em Reactive  control}: is the implementation  of tactics. This
  consists   in  moving,   targeting,  firing,   fleeing,  hit-and-run
  techniques  (also  knows  as  ``kiting'')  during  battle.  Reactive
  control focuses on a specific
  unit. %It is soldiers applying officers' orders.
%\end{itemize}
%
%\noindent
%Both  strategy  and  tactics  are  usually  called
%\textit{macro-management}, whereas reactive control is also called
%\textit{micro-management}. In addition of the three tasks shown above,
%one could also consider these two tasks that an AI has to deal with:
%\begin{itemize}
\item  {\em Terrain  analysis}: consists  in the  analysis of  regions
  composing the map: choke-points,  minerals and gas emplacements, low
  and high walkable grounds, islands, etc.
\item  {\em   Intelligence  gathering}:  corresponds   to  information
  collected  about the  opponent. Because  of the  fog-of-war, players
  must regularly send scouts to localize and spy enemy
  bases.%, and must see or infer what buildings and technology the enemy has developed in order to adapt its strategy.
\end{itemize}

@@

@@

\begin{figure}
    \centering
%    \includegraphics[width=0.7\columnwidth]{figures/levels_abstraction.pdf}
    \includegraphics[width=0.7\columnwidth]{figures/categories}
    \caption{RTS  AI  levels  of abstraction  and  typical sub-problems associated with them:
      %uncertainty  (coming  from  partial  observation  and  from  not knowing the  intentions of  the opponent)  is higher  for higher abstraction  levels.  
      Timings correspond  to  an
      estimate   of   the   duration   of   a   behavior   switch   in
      StarCraft. %Spatial and temporal  reasoning are indicated for the levels at which greedy solutions are not enough.
      }
    \label{fig:levels-abstraction}
\end{figure}

@@

\subsection{Strategy}
% Every AI needs to observe the world  in order to analyze it and take
% decisions. In  RTS games this  task isn't  trivial since we  have to
% deal with  imperfect information.   The most common  decision making
% are Decision  Trees. Some  of the advantages  of the  decision trees
% are:  simple to  understand and  interpret acting  like a  white box
% where  we  can  explain  why  a decision  is  taken,  and  they  are
% compatible with other decision  techniques. For complex decisions we
% have many algorithms  to generate an optimum decision  tree like ID3
% or C4.5, depends on the problem we will use one or another.

@@

% HARD-CODED:
\subsubsection{Hard-coded  approaches}
They have  been extensively  used in commercial  RTS games.   The most
common      ones     use      finite     state      machines     (FSM)
\cite{FSM_AIGameProgWisdom2003}  in   order  to  let  the   AI  author
hard-code the strategy  that the AI will employ. The  idea behind FSMs
is to decompose the AI behavior into easily manageable states, such as
``attacking'', ``gathering resources''  or ``repairing'' and establish
the  conditions that  trigger  transitions  between them.   Commercial
approaches also include Hierarchical FSMs,  in which FSMs are composed
hierarchically.    These  hard-coded   approaches   have  achieved   a
significant  amount of  success,  and, have  also been  used in  many academic  RTS AI
research systems.   However, these  hard-coded approaches  struggle to
encode  dynamic, adaptive  behaviors,  and are  easily exploitable  by
adaptive opponents.

@@

% PLANNING:
\subsubsection{Planning}
Approaches using  planning techniques have  also been explored  in the
literature.  For example  Onta\~{n}\'{o}n  et al.  \cite{CBR_Planning}
explored the use of real-time  case-based planning (CBP) in the domain
of  Wargus (a  Warcraft  II clone).  In their  work,  they used  human
demonstration to learn  plans, which are then composed  at run-time in
order  to   form  full-fledged  strategies   to  play  the   game.  In
\cite{PlanRetrieval} they improve over  their previous CBP approach by
using situation assessment for improving the quality and speed of plan
retrieval.  Hierarchical Task-Network  (HTN)  planning  has also  been
explored  with some  success in  the context  of simpler  first-person
shooter  games  \cite{HTNPlanning}.  Planning  approaches  offer  more
adaptivity    of   the    AI   strategy    compared   to    hard-coded
approaches. However, the real-time constraints  of RTS games limit the
planning approaches that  can be applied, HTN  and case-based planning
being  the  only  ones  explored  so  far.  Moreover,  none  of  these
approaches addresses any timing or scheduling issues, which are key in
RTS games.  On notable  exception is  the work  of Churchill  and Buro
\cite{churchill2011build}, who used planning in order to construct its
economic build-orders,  taking into account timing  constraints of the
different actions.

% MACHINE LEARNING:
\subsubsection{Machine Learning}
Concerning  machine   learning-based  approaches,  Weber   and  Mateas
\cite{WeberCig09}  proposed   a  data  mining  approach   to  strategy
prediction  and performed  supervised  learning  on labeled  StarCraft
replays.  Dereszynski et  al. \cite{HMMstrat_RTS_AIIDE11}  used Hidden
Markov Models (HMM) to learn the transition probabilities of sequences
of building  construction orders  and kept the  most probable  ones to
produce  probabilistic behavior  models (in  StarCraft). Synnaeve  and
Bessi\`{e}re   \cite{SynnaeveOpeningCig11}   used   the   dataset   of
\cite{WeberCig09} and  presented a  Bayesian semi-supervised  model to
learn from replays  and predict openings (early  game strategies) from
StarCraft  replays.   The  openings  are  labeled   by  EM  clustering
considering  appropriate  features. Then,  in  \cite{SynnaeveAIIDE11},
they presented  an unsupervised learning Bayesian  model for tech-tree
prediction, still using
replays. % Finally \cite{GemineImitative} used the same ``build-order features'' approach to learn to adapt the bot's build order to the opponents strategy in StarCraft 2.
Finally, evolutionary approaches to determine priorities of high level
tasks  were  explored  by  Young  and Hawes  in  their  QUORUM  system
\cite{young2012evolutionary},   showing    improvement   over   static
priorities.

% CBR:
\subsubsection{Case-Based Reasoning}
Also falling  into the machine-learning category,  a significant group
of    researchers   has    explored    case-based   reasoning    (CBR)
\cite{Aamodt94CBR}  approaches  for  strategic  decision  making.  For
example  Aha  et al.  \cite{LTW}  used  CBR  to perform  dynamic  plan
retrieval in  the Wargus domain.  Hsieh and Sun  \cite{HsiehS08} based
their work  on Aha et  al.'s CBR  model \cite{LTW} and  used StarCraft
replays   to  construct   states  and   building  sequences   (``build
orders''). Schadd et  al. \cite{SchaddBS07} applied a  CBR approach to
opponent  modeling through  hierarchically  structured  models of  the
opponent behavior and  they applied their work to the  Spring RTS game
(a ``Total Annihilation'' clone).  Jaidee et al. \cite{jaidee2011case}
study the  use of CBR for  automatic goal selection, while  playing an
RTS game. These goals will then determine which Q-tables to be used in
a  reinforcement  learning  framework.  Finally,  \v{C}ertick\'{y}  et
al. \cite{certicky2013cbr} used CBR to  build their army, based on the
opponent's army composition, and they pointed out on the importance of
proper scouting for better results.

% SCOUTING:
\subsubsection{Scouting}
One  final consideration  concerning strategy  is that  RTS games  are
typically  partially observable.  Games like  StarCraft implement  the
``fog-of-war'' idea, which basically means  that a player can only see
the areas  of the map close  to her own  units. Areas of the  map away
from the field of view of individual units are not observable. Players
need  to scout  in order  to obtain  information about  the opponent's
strategy. The size of the  state space in StarCraft prevents solutions
based on  POMDPs from being directly  applicable, and very few  of the
previous approaches deal  with this problem. Much work in  RTS game AI
assumes perfect information all the time.  For example, in the case of
commercial games, most AI implementations  cheat, since the AI can see
the  complete game  map  at all  times, while  the  human player  does
not. In order to  make the human player believe the  AI of these games
does not  cheat, sometimes  they simulate some  scouting tasks  as Bob
Fitch  described  in his  AIIDE  2011  keynote  for the  WarCraft  and
StarCraft game series.  Even if the StarCraft  AI competition enforces
fog-of-war, which  means that  bots are forced  to work  under partial
information, little published research exists on this topic. A notable
exception is the work of Weber  et al. \cite{WeberAIIDE11}, who used a
particle model with a linear trajectory update to track opponent units
under  fog-of-war  in StarCraft.  They  also  produced tactical  goals
through     reactive     planning     and     goal-driven     autonomy
\cite{WeberCig10,Weber10}, finding the more  relevant goal(s) to spawn
in unforeseen situations.

\subsection{Tactics}

@@

% MACHINE LEARNING:
\subsubsection{Machine Learning}
Concerning tactical  decision making,  many different  approaches have
been explored  such as machine  learning or game tree  search.  Hladky
and  Bulitko \cite{Hladky2008}  benchmarked hidden  semi-Markov models
(HSMM)  and particle  filters for  unit tracking.  Although they  used
first-person  shooter  (FPS)  games  for  their  experimentation,  the
results apply to  RTS games as well. They showed  that the accuracy of
occupancy maps  was improved using  movement models (learned  from the
player behavior)  in HSMM. Kabanza  et al. \cite{OBRecog}  improve the
probabilistic  hostile  agent  task  tracker  (PHATT  \cite{PHATT},  a
simulated HMM  for plan  recognition) by  encoding strategies  as HTN,
used   for   plan   and    intent   recognition   to   find   tactical
opportunities.   Sharma  et   al.  \cite{CBR-RL}   combined  CBR   and
reinforcement   learning   to   enable    reuse   of   tactical   plan
components. Cadena and Garrido  \cite{CadenaG11} used fuzzy CBR (fuzzy
case       matching)       for      strategic       and       tactical
planning.  \cite{SynnaeveTactics}  combined   space  abstraction  into
regions   from  \cite{Perkins10}   and  tactical-decision   making  by
assigning scores  (economical, defenses, etc.) to  regions and looking
for their  correspondences to  tactical moves (attacks)  in pro-gamers
replays.  Finally,  Miles \cite{miles2006co} created the  idea of {\em
  IMTrees}, a tree where each leaf  node is an influence map, and each
intermediate node  is a  combination operation  (sum, multiplication);
Miles used evolutionary algorithms to learn IMTrees for each strategic
decision in the game involving spatial reasoning by combining a set of
basic influence maps.

% GAME TREE SEARCH:
\subsubsection{Game  tree search}
These  techniques  have  also  been  explored  for  tactical  decision
making. Churchill and Buro \cite{churchill2012fast} presented the ABCD
algorithm  (Alpha-Beta  Considering  Durations), a  game  tree  search
algorithm   for   tactical   battles   in   RTS   games.    Chung   et
al. \cite{Chung05} applied Monte-Carlo  planning to a capture-the-flag
version  of Open  RTS.   Balla  and Fern  \cite{UCT}  applied the  UCT
algorithm (a  Monte Carlo Tree  Search algorithm) to  tactical assault
planning in Wargus. To make game tree search applicable at this level,
abstract game  state representations are  used in order to  reduce the
complexity. Uriarte and Onta\~{n}\'{o}n \cite{uriarte2014high,uriarte2014game} explored different game state abstractions in the context of Monte-Carlo Tree Search for high-level tactical reasoning in StarCraft. Other algorithms, such as Greedy Portfolio Search \cite{churchill2013portfolio}, perform abstraction at the level of actions, by employing a collection of predefined ``scripts'', and using these scripts as the possible actions that the players can execute in the context of game tree search.

@@

\subsection{Reactive Control}

>>>>

\subsubsection*{Planning}
As mentioned above, the  size of the state space in  RTS games is much
larger  than  that  of  traditional  board  games  such  as  Chess  or
Go. Additionally,  the number  of actions  that can  be executed  at a
given instant of time is  also much larger. Thus, standard adversarial
planning  approaches,  such  as  game tree  search  are  not  directly
applicable. As we  elaborate later, planning in RTS games  can be seen
as having multiple  levels of abstraction: at a  higher level, players
need long-term  planning capabilities,  in order  to develop  a strong
economy in the game; at a low level, individual units need to be moved
in coordination to  fight battles taking into account  the terrain and
the  opponent.  Techniques  that  can  address  these  large  planning
problems by either sampling, or  hierarchical decomposition do not yet
exist.

\subsubsection*{Learning}
Given  the  difficulties  in  playing  by  directly  using
adversarial  planning techniques,  many  research  groups have  turned
attention to  learning techniques. We  can distinguish three  types of
learning problems in RTS games:
\begin{itemize}
\item {\em Prior learning}: How can we exploit available data, such as
  existing replays,  or information  about specific maps  for learning
  appropriate strategies before hand? A significant amount of work has
  gone in this
  direction.%, such as \cite{WeberCIG09,SynnaeveCIG11,SynnaeveAIIDE11,OntanonMSR10}.
\item  {\em In-game  learning}: How  can bots  deploy online  learning
  techniques that allow them to  improve their game play while playing
  a  game?  These  techniques  might  include  reinforcement  learning
  techniques, but  also opponent modeling.  The main problem  again is
  the fact  that the state  space is too large  and the fact  that RTS
  games are partially observable.
\item {\em  Inter-game learning}:  What can be  learned from  one game
  that can  be used  to increase  the chances of  victory in  the next
  game? Some work has used simple game-theoretical solutions to select
  amongst a  pool of  predefined strategies,  but the  general problem
  remains unsolved.
\end{itemize}

@@

\subsubsection*{Uncertainty}
Adversarial planning under  uncertainty in domains of the  size of RTS
games is  still an unsolved  challenge.  In  RTS games, there  are two
main kinds  of uncertainty. First,  the game is  partially observable,
and players  cannot observe the  whole game  map (like in  Chess), but
need to scout in order to see what the opponent is doing. This type of
uncertainty   can  be   lowered  by   good  scouting,   and  knowledge
representation  (to  infer  what  is  possible  given  what  has  been
seen). However, scouting also means deliberately reducing economic progress in order to obtain information.
Second,  there is also  uncertainty arising from the  fact that
the games  are adversarial,  and a player  cannot predict  the actions
that the opponent(s)  will execute. For this type  of uncertainty, the
AI, as the human  player, can only build a sensible  model of what the
opponent is likely to do.
% [Santi:  I  commented  out  this  part,  since  I'm  not  sure  it's
% correct.  The second  type  of  uncertainty can  also  be linked  to
% data.  For example,  by analyzing  player replays,  one could  learn
% "plan recognition"  probability tables, which  can be used  to infer
% the strategy  of the opponent,  and thus, reduce the  uncertainty of
% the opponent's future  actions] Both these kinds  of uncertainty can
% be studied through probabilities, but only the first type, $P(Hidden
% | Seen)$, can be {\em easily} linked to
% data. %now for the clumsy part:
% One could  argue that the  extensional data  is the only  thing that
% matters because  it is the only  thing that has an  influence on the
% game, but in  reality, for very highly  skilled players, discovering
% the intentions of the  enemy is the key to winning.  It allows for a
% compact understanding of the game  state and development, while also
% giving additional information about the  future. Whereas a player is
% doing action  A to  follow it  up by action  B or  C depends  on its
% intention,  while the  information  about action  A  is visible  and
% information about  B or  C is  hidden, only  intention allows  for a
% choice.

\subsubsection*{Spatial and Temporal Reasoning}
% TODO just a paragraph on exposing why there are temporal and spatial
% reasoning in RTS AI.
Spatial reasoning is  related to each aspect  of terrain exploitation.
It  is  involved   in  tasks  such  as  building   placement  or  base
expansion.  In the  former,  the player  needs  to carefully  consider
building  positioning into  its  own  bases to  both  protect them  by
creating  defenses or walls  against invasions  and to  avoid bad  configurations
where large units could be stuck. In base expansion, the player has to
choose good available locations to build a new base, regarding its own
position and  opponent's bases. Finally,  spatial reasoning is  key to
tactical reasoning:  players need to  decide where to place  units for
battle, favoring, for instance,  engagements when the opponent's units
are lead into a bottleneck.

@@

\subsubsection*{Domain Knowledge Exploitation}
In traditional board  games such as Chess,  researchers have exploited
the  large  amounts  of  existing  domain  knowledge  to  create  good
evaluation  functions  to be  used  by  alpha-beta search  algorithms,
extensive opening books, or end-game tables. In the case of RTS games,
it  is still  unclear how  the  significantly large  amount of  domain
knowledge  (in the  forms or  strategy guides,  replays, etc.)  can be
exploited. Most  work in  this area has  focused on  two main
directions: (1) hard-coding existing strategies into bots (so that  bots only  need to
decide  which strategies  to deploy,  instead of  having to  solve the
complete  planning problem), and (2) mining large datasets
of  replays \cite{WeberCig09,synnaeve2012dataset} to automatically extract
 strategies,  trends   or  plans. % However,  StarCraft  games  are  quite  complex,  and  how  to
%automatically learn from such datasets is still an open problem.

\subsubsection*{Task Decomposition}
Most existing approaches  to play RTS games work by decomposing the  problem into   a    collection   of    smaller   problems to    be   solved
independently. Specifically, a common subdivision is:
\begin{itemize}
\item {\em  Strategy}: corresponds  to the high-level  decision making
  process.  This is  the highest  level  of abstraction  for the  game
  comprehension.  Finding an  efficient  strategy or  counter-strategy
  against a given opponent is key in RTS games, and concerns the whole
  set of units a player
  owns. %If we may allow ourselves a military analogy, strategy is orders given by generals at the HQ.
\item   {\em  Tactics}:   are  the   implementation  of   the  current
  strategy.  It  implies  army and  building  positioning,  movements,
  timing, and so on. Tactics concerns a group of
  units. %They are orders given by officers on the battlefield.
\item {\em Reactive  control}: is the implementation  of tactics. This
  consists   in  moving,   targeting,  firing,   fleeing,  hit-and-run
  techniques  (also  knows  as  ``kiting'')  during  battle.  Reactive
  control focuses on a specific
  unit. %It is soldiers applying officers' orders.
%\end{itemize}
%
%\noindent
%Both  strategy  and  tactics  are  usually  called
%\textit{macro-management}, whereas reactive control is also called
%\textit{micro-management}. In addition of the three tasks shown above,
%one could also consider these two tasks that an AI has to deal with:
%\begin{itemize}
\item  {\em Terrain  analysis}: consists  in the  analysis of  regions
  composing the map: choke-points,  minerals and gas emplacements, low
  and high walkable grounds, islands, etc.
\item  {\em   Intelligence  gathering}:  corresponds   to  information
  collected  about the  opponent. Because  of the  fog-of-war, players
  must regularly send scouts to localize and spy enemy
  bases.%, and must see or infer what buildings and technology the enemy has developed in order to adapt its strategy.
\end{itemize}

@@

\begin{figure}
    \centering
%    \includegraphics[width=0.7\columnwidth]{figures/levels_abstraction.pdf}
    \includegraphics[width=0.7\columnwidth]{categories}
    \caption{RTS  AI  levels  of abstraction  and  typical sub-problems associated with them:
      %uncertainty  (coming  from  partial  observation  and  from  not knowing the  intentions of  the opponent)  is higher  for higher abstraction  levels.  
      Timings correspond  to  an
      estimate   of   the   duration   of   a   behavior   switch   in
      StarCraft. %Spatial and temporal  reasoning are indicated for the levels at which greedy solutions are not enough.
      }
    \label{fig:levels-abstraction}
\end{figure}

@@

@@

\subsection*{Strategy}
% Every AI needs to observe the world  in order to analyze it and take
% decisions. In  RTS games this  task isn't  trivial since we  have to
% deal with  imperfect information.   The most common  decision making
% are Decision  Trees. Some  of the advantages  of the  decision trees
% are:  simple to  understand and  interpret acting  like a  white box
% where  we  can  explain  why  a decision  is  taken,  and  they  are
% compatible with other decision  techniques. For complex decisions we
% have many algorithms  to generate an optimum decision  tree like ID3
% or C4.5, depends on the problem we will use one or another.

@@

% HARD-CODED:
\subsubsection*{Hard-coded  approaches}
They have  been extensively  used in commercial  RTS games.   The most
common      ones     use      finite     state      machines     (FSM)
\cite{FSM_AIGameProgWisdom2003}  in   order  to  let  the   AI  author
hard-code the strategy  that the AI will employ. The  idea behind FSMs
is to decompose the AI behavior into easily manageable states, such as
``attacking'', ``gathering resources''  or ``repairing'' and establish
the  conditions that  trigger  transitions  between them.   Commercial
approaches also include Hierarchical FSMs,  in which FSMs are composed
hierarchically.    These  hard-coded   approaches   have  achieved   a
significant  amount of  success,  and, have  also been  used in  many academic  RTS AI
research systems.   However, these  hard-coded approaches  struggle to
encode  dynamic, adaptive  behaviors,  and are  easily exploitable  by
adaptive opponents.

@@

% PLANNING:
\subsubsection*{Planning}
Approaches using  planning techniques have  also been explored  in the
literature.  For example  Onta\~{n}\'{o}n  et al.  \cite{CBR_Planning}
explored the use of real-time  case-based planning (CBP) in the domain
of  Wargus (a  Warcraft  II clone).  In their  work,  they used  human
demonstration to learn  plans, which are then composed  at run-time in
order  to   form  full-fledged  strategies   to  play  the   game.  In
\cite{PlanRetrieval} they improve over  their previous CBP approach by
using situation assessment for improving the quality and speed of plan
retrieval.  Hierarchical Task-Network  (HTN)  planning  has also  been
explored  with some  success in  the context  of simpler  first-person
shooter  games  \cite{HTNPlanning}.  Planning  approaches  offer  more
adaptivity    of   the    AI   strategy    compared   to    hard-coded
approaches. However, the real-time constraints  of RTS games limit the
planning approaches that  can be applied, HTN  and case-based planning
being  the  only  ones  explored  so  far.  Moreover,  none  of  these
approaches addresses any timing or scheduling issues, which are key in
RTS games.  On notable  exception is  the work  of Churchill  and Buro
\cite{churchill2011build}, who used planning in order to construct its
economic build-orders,  taking into account timing  constraints of the
different actions.

% MACHINE LEARNING:
\subsubsection*{Machine Learning}
Concerning  machine   learning-based  approaches,  Weber   and  Mateas
\cite{WeberCig09}  proposed   a  data  mining  approach   to  strategy
prediction  and performed  supervised  learning  on labeled  StarCraft
replays.  Dereszynski et  al. \cite{HMMstrat_RTS_AIIDE11}  used Hidden
Markov Models (HMM) to learn the transition probabilities of sequences
of building  construction orders  and kept the  most probable  ones to
produce  probabilistic behavior  models (in  StarCraft). Synnaeve  and
Bessi\`{e}re   \cite{SynnaeveOpeningCig11}   used   the   dataset   of
\cite{WeberCig09} and  presented a  Bayesian semi-supervised  model to
learn from replays  and predict openings (early  game strategies) from
StarCraft  replays.   The  openings  are  labeled   by  EM  clustering
considering  appropriate  features. Then,  in  \cite{SynnaeveAIIDE11},
they presented  an unsupervised learning Bayesian  model for tech-tree
prediction, still using
replays. % Finally \cite{GemineImitative} used the same ``build-order features'' approach to learn to adapt the bot's build order to the opponents strategy in StarCraft 2.
Finally, evolutionary approaches to determine priorities of high level
tasks  were  explored  by  Young  and Hawes  in  their  QUORUM  system
\cite{young2012evolutionary},   showing    improvement   over   static
priorities.

% CBR:
\subsubsection*{Case-Based Reasoning}
Also falling  into the machine-learning category,  a significant group
of    researchers   has    explored    case-based   reasoning    (CBR)
\cite{Aamodt94CBR}  approaches  for  strategic  decision  making.  For
example  Aha  et al.  \cite{LTW}  used  CBR  to perform  dynamic  plan
retrieval in  the Wargus domain.  Hsieh and Sun  \cite{HsiehS08} based
their work  on Aha et  al.'s CBR  model \cite{LTW} and  used StarCraft
replays   to  construct   states  and   building  sequences   (``build
orders''). Schadd et  al. \cite{SchaddBS07} applied a  CBR approach to
opponent  modeling through  hierarchically  structured  models of  the
opponent behavior and  they applied their work to the  Spring RTS game
(a ``Total Annihilation'' clone).  Jaidee et al. \cite{jaidee2011case}
study the  use of CBR for  automatic goal selection, while  playing an
RTS game. These goals will then determine which Q-tables to be used in
a  reinforcement  learning  framework.  Finally,  \v{C}ertick\'{y}  et
al. \cite{certicky2013cbr} used CBR to  build their army, based on the
opponent's army composition, and they pointed out on the importance of
proper scouting for better results.

% SCOUTING:
\subsubsection*{Scouting}
One  final consideration  concerning strategy  is that  RTS games  are
typically  partially observable.  Games like  StarCraft implement  the
``fog-of-war'' idea, which basically means  that a player can only see
the areas  of the map close  to her own  units. Areas of the  map away
from the field of view of individual units are not observable. Players
need  to scout  in order  to obtain  information about  the opponent's
strategy. The size of the  state space in StarCraft prevents solutions
based on  POMDPs from being directly  applicable, and very few  of the
previous approaches deal  with this problem. Much work in  RTS game AI
assumes perfect information all the time.  For example, in the case of
commercial games, most AI implementations  cheat, since the AI can see
the  complete game  map  at all  times, while  the  human player  does
not. In order to  make the human player believe the  AI of these games
does not  cheat, sometimes  they simulate some  scouting tasks  as Bob
Fitch  described  in his  AIIDE  2011  keynote  for the  WarCraft  and
StarCraft game series.  Even if the StarCraft  AI competition enforces
fog-of-war, which  means that  bots are forced  to work  under partial
information, little published research exists on this topic. A notable
exception is the work of Weber  et al. \cite{WeberAIIDE11}, who used a
particle model with a linear trajectory update to track opponent units
under  fog-of-war  in StarCraft.  They  also  produced tactical  goals
through     reactive     planning     and     goal-driven     autonomy
\cite{WeberCig10,Weber10}, finding the more  relevant goal(s) to spawn
in unforeseen situations.

\subsection*{Tactics}

@@

% MACHINE LEARNING:
\subsubsection*{Machine Learning}
Concerning tactical  decision making,  many different  approaches have
been explored  such as machine  learning or game tree  search.  Hladky
and  Bulitko \cite{Hladky2008}  benchmarked hidden  semi-Markov models
(HSMM)  and particle  filters for  unit tracking.  Although they  used
first-person  shooter  (FPS)  games  for  their  experimentation,  the
results apply to  RTS games as well. They showed  that the accuracy of
occupancy maps  was improved using  movement models (learned  from the
player behavior)  in HSMM. Kabanza  et al. \cite{OBRecog}  improve the
probabilistic  hostile  agent  task  tracker  (PHATT  \cite{PHATT},  a
simulated HMM  for plan  recognition) by  encoding strategies  as HTN,
used   for   plan   and    intent   recognition   to   find   tactical
opportunities.   Sharma  et   al.  \cite{CBR-RL}   combined  CBR   and
reinforcement   learning   to   enable    reuse   of   tactical   plan
components. Cadena and Garrido  \cite{CadenaG11} used fuzzy CBR (fuzzy
case       matching)       for      strategic       and       tactical
planning.  \cite{SynnaeveTactics}  combined   space  abstraction  into
regions   from  \cite{Perkins10}   and  tactical-decision   making  by
assigning scores  (economical, defenses, etc.) to  regions and looking
for their  correspondences to  tactical moves (attacks)  in pro-gamers
replays.  Finally,  Miles \cite{miles2006co} created the  idea of {\em
  IMTrees}, a tree where each leaf  node is an influence map, and each
intermediate node  is a  combination operation  (sum, multiplication);
Miles used evolutionary algorithms to learn IMTrees for each strategic
decision in the game involving spatial reasoning by combining a set of
basic influence maps.

% GAME TREE SEARCH:
\subsubsection*{Game  tree search}
These  techniques  have  also  been  explored  for  tactical  decision
making. Churchill and Buro \cite{churchill2012fast} presented the ABCD
algorithm  (Alpha-Beta  Considering  Durations), a  game  tree  search
algorithm   for   tactical   battles   in   RTS   games.    Chung   et
al. \cite{Chung05} applied Monte-Carlo  planning to a capture-the-flag
version  of Open  RTS.   Balla  and Fern  \cite{UCT}  applied the  UCT
algorithm (a  Monte Carlo Tree  Search algorithm) to  tactical assault
planning in Wargus. To make game tree search applicable at this level,
abstract game  state representations are  used in order to  reduce the
complexity. Uriarte and Onta\~{n}\'{o}n \cite{uriarte2014high,uriarte2014game} explored different game state abstractions in the context of Monte-Carlo Tree Search for high-level tactical reasoning in StarCraft. Other algorithms, such as Greedy Portfolio Search \cite{churchill2013portfolio}, perform abstraction at the level of actions, by employing a collection of predefined ``scripts'', and using these scripts as the possible actions that the players can execute in the context of game tree search.

@@

\subsection*{Reactive Control}
